---
title: "146 5.2 Dirilecht"
output: html_notebook
---

***Model the following social mobility data set using a multinomial likelihood with a Dirichlet prior. The data set contains the results of a survey on male workers, and specifically what type of work (farm work, unskilled laborer, skilled laborer, professional) a son does, depending the type of work his father does.***
```{r}
# clears workspace: 
rm(list=ls()) 
# load data:
data <- read.csv("/Users/oba2311/Desktop/Minerva/Junior/CS146/week 5/socialmobility.csv")
head(data)
```

***Set up your Dirichlet prior so that you have a uniform distribution over the probability vector parameter of the multinomial. Write R code to compute the posterior Dirichlet distribution, given your prior hyperparameters and the data.***
Load & configure RStan:
```{r}
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

```{r}

# We have to tell Stan what data to expect, what our parameters are and what
# the likelihood and prior are. Since the posterior is just proportional to
# the product of the likelihood and the prior, we don't distinguish between
# them explicitly in the model below. Every distribution we specify is
# automatically incorporated into the product.

#dataprep:

stan_model <- "

data {
  int<lower=1> patients;  // total number of patients in the studies.
  int<lower=0> improved;  // number of improved patients.
  
}

parameters {
  real<lower=1> alpha;    // prior parameters, alpha is 1 to get a uniform distribution.
  real<lower=0,upper=1> theta;  // probability of workig in fathre's proffession.
}

model {
  theta ~ dirichlet_lpdf(theta,alpha);        // prior over theta
  improved ~ multinomial(theta);  // likelihood function
}

"

# Fit the model to the data. This will generate samples over all parameters of our
# model. This will take a little while to run. Stan will feel slow even on simple models.

# NOTE: The cores=1 line is necessary to make stan work on CoCalc. Please remember
# to add it whenever you call the stan() function — the call will fail if you don't.

# NOTE: If you see a weird error message that looks like this, you can safely ignore it:
#
# #  define BOOST_NO_CXX11_RVALUE_REFERENCES
#  ^
# <command-line>:0:0: note: this is the location of the previous definition


fit <- stan(
    model_code = stan_model,
    data = data,
    cores = 2,
    save_warmup = FALSE
)
```


```{r}
# The commands below are useful for a quick overview:
print(samples)  # a rough summary
print(summary(samples))  # more detailed summary
plot(samples)   # a visual representation

```



Michael's code:
```{r}
library(gtools)
socialmobility <- read.csv("socialmobility.csv")
alphas = c(1, 1, 1, 1)

son_conditioned_on_father <- function (kind, alphas=c(1, 1, 1, 1)) {
  return(socialmobility[socialmobility$father == kind, 3] + alphas)
}

father_conditioned_on_son <- function (kind, alphas=c(1, 1, 1, 1)) {
  return(socialmobility[socialmobility$son == kind, 3] + alphas)
}

num_samples <- 10000000
quantile(rdirichlet(5, son_conditioned_on_father('unskilled'))[,3],
         probs=c(.025, .975))

quantile(rdirichlet(5, father_conditioned_on_son('professional'))[,1],
         probs=c(.025, .975))

```
Jones Code:
```{r}
library(gtools)

data = read.csv("socialmobility.csv")

# defining prior alphas

alphas = rep(1, each=16)

# getting posterior alpha function

Multino_Dirich_posterior <- function(prior_alpha_vector, data_vector){
  for (i in 1:length(prior_alpha_vector)) {
    prior_alpha_vector[i] = prior_alpha_vector[i] + data_vector[i]
  }
  return(prior_alpha_vector)
}

# data as numeric

data_vector <- as.numeric(data$count)

# getting prior alphas

new_alphas <- Multino_Dirich_posterior(alphas,data_vector)

# generating data

x <- rdirichlet(1000, new_alphas )

# unskilled father -> skilled son
quantile(x[,7],  probs = c(0.025, 0.975))

# proffesional son -> farm father
quantile(x[,4],  probs = c(0.025, 0.975))

```


